---
title: "Untitled"
author: "Team 15"
date: "2/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Loading packages and the toy horse data
rm (list = ls())
require("cluster")
require("fpc")
require("factoextra")
require("gridExtra")
library(cluster)
library(fpc)
library(factoextra)
library(gridExtra)
library(conjoint)

dir= '/Users/cynth1a/Desktop/Analytic Design/cases & assignment'
setwd(dir)
load("GBA424 - Toy Horse Case Data.rdata")

```


##### Part A.Getting Individual Part-Utilities using Linear Regression #####
```{r}
#data cleaning: drop rows with NA
conjointData1 = conjointData[complete.cases(conjointData), ]
conjointDataNA <- conjointData[is.na(conjointData$ratings),]

#create a dataframe and fill each individual's part-utilities
coef <- data.frame(ID = rep(NA, 200), intercept = rep(NA,200), 
                   price_119.99 = rep(NA,200), size_26 = rep(NA,200), motion_rock = rep(NA,200),
                   style_glam = rep(NA,200))
for(i in 1:200) {
  a = lm(ratings~price+size+motion+style, data = subset(conjointData1, ID==i))
  coef[i,1] = i
  coef[i,2:6] = a$coefficient[1:5]
  coef
}
```


```{r}
#predict for missing profiles 3,6,10,16
for (i in 1:200){
  for (j in c(3,6,10,16)){
    conjointDataNA[conjointDataNA$ID == i & conjointDataNA$profile == j, 3] = 
      coef[i,2] + 
      conjointDataNA[conjointDataNA$ID == i & conjointDataNA$profile == j, 4] * coef[i,3] + 
      conjointDataNA[conjointDataNA$ID == i & conjointDataNA$profile == j, 5] * coef[i,4] + 
      conjointDataNA[conjointDataNA$ID == i & conjointDataNA$profile == j, 6] * coef[i,5] + 
      conjointDataNA[conjointDataNA$ID == i & conjointDataNA$profile == j, 7] * coef[i,6]
  }
}
```


```{r}
#merge and reorder the dataframe by ID and profile
conjointData_full <-rbind(conjointDataNA, conjointData1)
conjointData_full <-conjointData_full[order(conjointData_full$ID, conjointData_full$profile),]
```



###### Part B.Benefit Segmentation via Cluster Analysis ######
```{r}
##Evaluate number of clusters to use on data with visualizations
clustTest = function(toClust,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
  if(scale){ toClust = scale(toClust);}
  set.seed(seed);   # set random number seed before doing cluster analysis
  wss <- (nrow(toClust)-1)*sum(apply(toClust,2,var))
  for (i in 2:maxClusts) wss[i] <- sum(kmeans(toClust,centers=i,nstart=nstart,iter.max=iter.max)$withinss)
  ##gpw essentially does the following plot using wss above. 
  #plot(1:maxClusts, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
  gpw = fviz_nbclust(toClust,kmeans,method="wss",iter.max=iter.max,nstart=nstart,k.max=maxClusts) #alternative way to get wss elbow chart.
  pm1 = pamk(toClust,scaling=TRUE)
  ## pm1$nc indicates the optimal number of clusters based on 
  ## lowest average silhoutte score (a measure of quality of clustering)
  #alternative way that presents it visually as well.
  gps = fviz_nbclust(toClust,kmeans,method="silhouette",iter.max=iter.max,nstart=nstart,k.max=maxClusts) 
  if(print){
    grid.arrange(gpw,gps, nrow = 1)
  }
  list(wss=wss,pm1=pm1$nc,gpw=gpw,gps=gps)
}
```


```{r}
## Runs a set of clusters as kmeans
runClusts = function(toClust,nClusts,print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100){
  if(length(nClusts)>4){
    warning("Using only first 4 elements of nClusts.")
  }
  kms=list(); ps=list();
  for(i in 1:length(nClusts)){
    kms[[i]] = kmeans(toClust,nClusts[i],iter.max = iter.max, nstart=nstart)
    ps[[i]] = fviz_cluster(kms[[i]], geom = "point", data = toClust) + ggtitle(paste("k =",nClusts[i]))
    
  }
  library(gridExtra)
  if(print){
    tmp = marrangeGrob(ps, nrow = 2,ncol=2)
    print(tmp)
  }
  list(kms=kms,ps=ps)
}
```


```{r}
##  pie chart with membership percentages
##  ellipse plot that indicates cluster definitions against principle components
##  barplot of the cluster means
plotClust = function(km,toClust,discPlot=FALSE){
  nc = length(km$size)
  if(discPlot){par(mfrow=c(2,2))}
  else {par(mfrow=c(3,1))}
  percsize = paste(1:nc," = ",format(km$size/sum(km$size)*100,digits=2),"%",sep="")
  pie(km$size,labels=percsize,col=1:nc)
  
  clusplot(toClust, km$cluster, color=TRUE, shade=TRUE,
           labels=2, lines=0,col.clus=1:nc); #plot clusters against principal components
  
  if(discPlot){
    plotcluster(toClust, km$cluster,col=km$cluster); #plot against discriminant functions ()
  }
  rng = range(km$centers)
  dist = rng[2]-rng[1]
  locs = km$centers+.05*dist*ifelse(km$centers>0,1,-1)
  bm = barplot(km$centers,beside=TRUE,col=1:nc,main="Cluster Means",ylim=rng+dist*c(-.1,.1))
  text(bm,locs,formatC(km$centers,format="f",digits=1))
}

partworth = coef[2:6]

checks = clustTest(partworth,print=TRUE,scale=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
clusts = runClusts(partworth,c(2,3,4,9),print=TRUE,maxClusts=15,seed=12345,nstart=20,iter.max=100)
plotClust(clusts[[1]][[2]],partworth)
```
We tried four different cluster schemes with 2 clusters, 3 clusters, 4 clusters, and 9 clusters. According to the two plots that show the optimal number of clusters, we can know that the optimal number of clusters is 3. Elbow rule suggests around 3 clusters, focal points at 3 clusters as well.
First segment (black)：care about the price most;prefer 26 inches ; prefer bouncing the most; prefer racing the most ---- Profile 4
Second segment (red): prefer low price; likes 18-inch one the most; prefer rocking the most; prefer glamorous ---- Profile 14
Third segment (green): care about the size most; they are the least price sensitive; prefer 26-inch the most; prefer rocking motion; prefer glamorous the most. ---- Profile 16



##### Part C. A Priori segmentation using Age and Gender #####
```{r}
# Create a dataframe to combine conjoint data with demographics data
conjointDemo <- merge(conjointData1, respondentData, by = 'ID')
# Without Segmentation
summary(lm(ratings~price+size+motion+style, data = conjointDemo))
```
Price is the most influential attribute; style is the least influential attribute
Importance: price > size > motion > style 
--- Profile 16


## A Priori Segmentation using age and gender (seperately) ##
```{r}
#age = 0: 2 years old
age0 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 0))
summary(age0)
```
For children whose age is 2 years, prefer the lower price ($119.99). Comparing to children who are 3-4 years old, they also care less about size and prefer 26-inches. They also prefer rocking motion and glamorous style.
--- Profile 16


```{r}
#age = 1: 3-4 years old
age1 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 1))
summary(age1)
```
For children whose age is around 3-4 years,  they care about style more than motion and they prefer bouncing motion more than rocking.
---- Profile 12

```{r}
#gender = 0: boy
gender0 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, gender == 0))
summary(gender0)
```
For boys, they care about style more than motion. They prefer bouncing motion and prefer racing tyle.
---- profile 4


```{r}
#gender = 1: girl
gender1 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, gender == 1))
summary(gender1)
```
For male and female, they all want lower price and 26-inches. However, comparing to female, male care more about price, female care more about size. they all care about style more than motion, but male prefer bouncing motion and racing style, while female prefer rocking motion and glamour style.


## A Priori Segmentation using age and gender (combined) ##
```{r}
#age = 0, gender = 0, 2 year-old boy 
age0gender0 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 0 & gender == 0))
summary(age0gender0)

#age = 0, gender = 1, 2 year-old girl 
age0gender1 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 0 & gender == 1))
summary(age0gender1)
```


```{r}
#age = 1, gender = 0, 3-4 year-old boy
age1gender0 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 1 & gender == 0))
summary(age1gender0)

#age = 1, gender = 1, 3-4 year-old girl
age1gender1 = lm(ratings~price+size+motion+style, data = subset(conjointDemo, age == 1 & gender == 1))
summary(age1gender1)
```
There are 4 segments using two demographics combined: younger boy (2), elder boy (3-4), younger girl (2), and elder girl (3-4).
In both age group, boys’ parents care more about size and price.
2 years old boy prefer bouncing motion while 3-4 years old boy prefer rock motion, because of their different playing habit.
Overall, profile 4, 8, 12,16 should be launched to target those segments, especially profile 16, which is prefered by 4 segments. 


```{r}
## Statistical Significance test
#base on gender
summary(lm(ratings~price+size+motion+style+gender:price+gender:size+gender:motion+gender:style, data = conjointDemo))
#base on age
summary(lm(ratings~price+size+motion+style+age:price+age:size+age:motion+age:style, data = conjointDemo))

```
- The interaction between price and gender is not statistically significant, which means that the difference between price’s effect on boy and girl groups’ utilities is not significant--- we may not care much about price’s impact on different gender groups.
- The interaction effect between age and price, as well as between style and age is statistically insignificant.
- Compared to age, gender is a more significant segmentation factor according to the P-value of interacting terms (significant results on three out of four attributes).


```{r}
# Create a dataframe and fill in the summary of part-utilities of different segments within A Priori Segmentation
coefDemo <- data.frame(demo = rep(NA, 8), intercept = rep(NA,8), 
                   price_119.99 = rep(NA,8), size_26 = rep(NA,8), motion_rock = rep(NA,8),
                   style_glam = rep(NA,8))
rownames(coefDemo) = c('age0','age1','gender0','gender1','age0gender0','age0gender1','age1gender0','age1gender1')
coefDemo[1,2:6] = age0$coefficients
coefDemo[2,2:6] = age1$coefficients
coefDemo[3,2:6] = gender0$coefficients
coefDemo[4,2:6] = gender1$coefficients
coefDemo[5,2:6] = age0gender0$coefficients
coefDemo[6,2:6] = age0gender1$coefficients
coefDemo[7,2:6] = age1gender0$coefficients
coefDemo[8,2:6] = age1gender1$coefficients
coefDemo[,1] = rownames(coefDemo)

coefDemo
```


##### Part D. Market Simulations and Profitability Calculations #####
```{r}
## Calculate Market Share based on First choice (superficially)
choice = data.frame(ID = rep(NA,200), profile = rep(NA,200), rating = rep(NA,200),  price = rep(NA,200), size = rep(NA,200), motion = rep(NA,200),
               style = rep(NA,200))
for(i in 1:200) {
  Y = subset(conjointData_full, ID==i)[,3]
  choice[i,] = subset(conjointData_full, ID==i)[which.max(Y),]
}

# Get profiles that are first choice
unique(choice$profile) #choices include profile 16,6,12,14,4,2,8,15,13
choiceProfile = c(16,6,12,14,4,2,8,15,13)

mktShare = data.frame()
for (z in 1:length(choiceProfile)) {
  marketshare = nrow(subset(choice, profile == choiceProfile[z]))/nrow(choice)
  mktShare[z,1] = choiceProfile[z]
  mktShare[z,2] = marketshare
}
colnames(mktShare) = c('profile', "market_share")
mktShare <- mktShare[order(mktShare$market_share, decreasing = T),]
mktShare

```
We first calculated market shares superficially: and we found out that the top three profiles are 16,4, and 14. Then, we would like to use simulations to further look into market shares.


```{r}
##calculate market share using market simulation tools
# Rate Chart with each profile as a column
Rate = as.data.frame(c(1:200))

for (i in 1:200){
  ID_data = conjointData_full[conjointData_full$ID==i,]
  Rate[i,c(2:17)] = t(ID_data[,3])
}

colnames(Rate) <- c('ID','P1','P2','P3','P4',
                    'P5','P6','P7','P8','P9',
                    'P10','P11','P12','P13','P14',
                    'P15','P16')
RateData <- Rate[,-1]
```



```{r}
# Simulated market share function
simFCSharesA = function(scen,data,ascend=TRUE){ 
  inmkt = data[,scen] #construct the subsetted matrix of options
  scores = matrix(c(rep(0,length(scen))))
  if(ascend){ #if ranks 1 is best
    for (i in seq(1:nrow(inmkt))) {
      z = min(inmkt[i,])
      if (sum(inmkt[i,]==z) == 1) {
        scores[inmkt[i,]==z] = scores[inmkt[i,]==z] + 1
      }
      else {
        j = sum(inmkt[i,]==z)
        scores[inmkt[i,]==z] = scores[inmkt[i,]==z] + 1/j
      }
    }
  } else { #else the best rank is the largest number
    for (i in seq(1:nrow(inmkt))) {
      y = max(inmkt[i,])
      if (sum(inmkt[i,]==y) == 1) {
        scores[inmkt[i,]==y] = scores[inmkt[i,]==y] + 1
      }
      else {
        j = sum(inmkt[i,]==y)
        scores[inmkt[i,]==y] = scores[inmkt[i,]==y] + 1/j
      }
    }
  }
  marketShare = data.frame(t(scores/i))
  colnames(marketShare) = colnames(inmkt)
  marketShare
}
```


```{r}
#create scenarios (add products with higher market shares and use different conbination plans)
scen0 = c(13, 5, 7) #current market
scen1 =c(13, 5, 4, 14, 16, 8)  #launch top 3 profiles
scen2 = c(4,13,5,8)  #delete 14, 16 (lowest market share), launch top 3 profiles
scen3 = c(4,14,5,8) 
scen4 = c(4,5,8) #delete 14
scen5 = c(4,5,12,8) #add 12 
scen6 = c(4,12,8) # drop 5
scen7 = c(12,5,8) # replace 4 with 5

#market shares 
simFCSharesA(scen0, RateData[])
simFCSharesA(scen1, RateData[])
simFCSharesA(scen2, RateData[])
simFCSharesA(scen3, RateData[])
simFCSharesA(scen4, RateData[])
simFCSharesA(scen5, RateData[])
simFCSharesA(scen6, RateData[])
simFCSharesA(scen7, RateData[])
```
Here are our key assumptiona when simulating different scenarios. First, when Early Riders keep our status quo (Profiles 5 and 13), the only competitor would keep his status quo (Profile 7); when Early Riders add one or more new product line, the competitor would take the action of lowering retail price to $119.99 (which is Profile 8). Second, we assume that Early Riders do not offer Profiles 7 and 8, which are only  provided by the competitor in the market.
We add Profiles 4, 14, and 16 because the cluster analysis in Part B suggest that the three segments prefer Profiles 4, 14 and 16 respectively. 
We add Profile 12 into consideration because in the A Priori Segmentation analysis in Part C, the preferred profiles according to age and gender (respectively) are Profiles 4, 16, 12, and 8. Since we have already considered Profiles 4 and 16 from Part B and only competitor is offering Profile 8, we add Profile 12 for a more comprehensive analysis.


```{r}
##Costs = 20,000/year * #products + $20,000/3 *#products not in existing set
#Shortrun Simulated Profits - 1 year period
Profit0 = 0.385*78.99*4000+0.42*78.99*4000-20000*2 # $214,348
Profit1 = 0.395*78.99*4000+0.42*78.99*4000+0.185*66.99*4000-20000*5-20000/3*3 # $173,747
Profit2 = 0.185*66.99*4000+0.395*78.99*4000+0.42*78.99*4000-20000*3-20000/3*1 # $240,413
profit3 = 0.215*66.99*4000+0.005*62.99*4000+0.78*4000*78.99-20000*3-20000/3*2 # $231,987
profit4 = 0.215*4000*66.99+0.785*4000*78.99-20000*2-20000/3*1 # $258,973
profit5 = 0.15*4000*66.99+0.755*4000*78.99+0.095*4000*66.99-20000*3-20000/3*2 # $230,867
profit6 = 0.45*4000*(95.99-29)+0.265*4000*(95.99-29)-20000*2-20000/3*2 # $138,258
profit7 = 0.79*4000*(111.99-33)+0.20*4000*(95.99-29)-20000*2-20000/3*1 # $256,534
```


```{r}
#Longrun Simulated Profits - 5 years period
Profit0.long = 5*(0.385*78.99*4000+0.42*78.99*4000-20000*2) #1071739
Profit1.long = 5*(0.395*78.99*4000+0.42*78.99*4000+0.185*66.99*4000-20000*5)-20000/3*3 #1015400
Profit2.long = 5*(0.185*66.99*4000+0.395*78.99*4000+0.42*78.99*4000-20000*3)-20000/3*1 #1228733
profit3.long = 5*(0.215*66.99*4000+0.005*62.99*4000+0.78*4000*78.99-20000*3)-20000/3*2 #1213267
profit4.long = 5*(0.215*4000*66.99+0.785*4000*78.99-20000*2)-20000/3*1 #1321533
profit5.long = 5*(0.15*4000*66.99+0.755*4000*78.99+0.095*4000*66.99-20000*3)-20000/3*2 #1207667
profit6.long = 5*(0.45*4000*(95.99-29)+0.265*4000*(95.99-29)-20000*2)-20000/3*2 #744623.7
profit7.long = 5*(0.79*4000*(111.99-33)+0.20*4000*(95.99-29)-20000*2)-20000/3*1 #1309335

```
Scenario 4 (with Profiles 4, 5, and 8) has highest profit for both longrun and shortrun situation. We hope that with our analysis, our client, Early Riders, would be able to better segment,target and positiontheir toy horses in the market.
